Redundant calibration as a complex optimization problem

### 2017-01-05 paper.tex
worked a bit on the introduction
added introduction to paper

\begin{abstract}
Traditionally redundant calibration is implemented via a hierarchical algorithmic strategy, i.e. a basic \textsc{firstcal} is performed, followed by \textsc{logcal} and \textsc{lincal}.  
\textsc{logcal} requires the use of the logartihm, while \textsc{lincal} makes use of the gradient-based Gauss-Newton algorithm.
We show that instead of splitting our calibration problem into its phase and amplitude components,
which is the approach taken by \textsc{lincal}, we can treat the complex unknown model parameters and their complex conjugates as separate unknowns instead. 
We further propose the use of the Levenberg-Marquardt algorithm, instead of the Gauss-Newton algorithm, to improve the convergence properties of the proposed complex
framework, which in turn reduces the number of layers that redundant calibration pipelines require. We also compare two previously proposed methods to speedup redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method. We found that the alternating 
direction implicit method outperforms the preconditioned conjugate gradient method, however the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases. 
We mainly focus on the complex framework in this paper rather than the amplitude and phase approach as it makes it trivial to relate the alternating direction implicit method 
and the Levenberg-Marquardt algorithm with each other.
\end{abstract}

@@

Redundant calibration was already in use in the 1980's. \citet{Noordam1982} used redundant calibration and the WSRT (Westerbork Synthesis Radio Telescope) to produce high-dynamic range images of 3C84.
The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation, redundant 
calibration is transformed into a linear problem \citep{Wieringa1992,Camps2003,Liu2010}. This approach is commenly referred to as \textsc{logcal}.
We briefly discuss \textsc{logcal} in Appendix~\ref{sec:logcal}. The next improvement in redundant calibration was to perform a firt-order taylor expansion, which is equivalent to the Gauss-Newton (GN) algorithm \citep{Kurien2016}, of the measurement equation an then to solve for the
perturbations of the model-parameters \citep{Liu2010}. The implementation of  this approach is called \textsc{lincal}. \textsc{lincal} is derived from first principals in Appendix~\ref{sec:lincal}. 
\textsc{logcal} and \textsc{lincal} are both implemented in \textsc{omnical}\footnote{https://github.com/jeffzhen/omnical}, which is a layered redundant calibration pipeline \citep{Zheng2014,Ali2015}.
The solutions of \textsc{logcal} are used as initial input for \textsc{lincal}, since the GN algorithm requires that its starting parameter guesses be in the vacinity 
of the solution.

Two ways of speeding up redundant calibration have also been proposed recently. The first technique makes use of the ADI method, in which we alternate between solving the gains, their conjugates and 
the true observed visibilities \citep{Wijnholds2012,Marthi2014}. The second is to use the preconditioned conjugate gradthient (PCG) method. PCG allows us to exploit the sparsity inherent in the redundant calibration problem, 
to speed it up \citep{Liu2010}. A proper comparison between these two approaches has not been done before.

Newly built instruments like LOFAR, ..... has also sparked a renewed interest in redundant calibration. Redundancy is also advantageous for detecting the EoR,
which explains why some of the previsously mentioned instruments have redundant layouts.

Layered approach, LOGCAL, \textsc{lincal}, Stefan's conference paper, Noordam, other ASTRON paper, newest papers on redundant calibration, marthi paper etc...

EOR SECTION SOMEWHERE HERE

MAIN CONTRIBUTION OF THIS PAPER COMES HERE

>>>>

\begin{abstract}
Traditionally redundant calibration is implemented via a hierarchical algorithmic strategy, i.e. a basic \textsc{firstcal} is performed, followed by \textsc{logcal} and \textsc{lincal}.  
\textsc{logcal} requires the use of the logartihm, while \textsc{lincal} makes use of the gradient-based Gauss-Newton algorithm.
We show that instead of splitting our calibration problem into its phase and amplitude components,
which is the approach taken by \textsc{lincal}, we can treat the complex unknown model parameters and their complex conjugates as separate unknowns instead. 
We further propose the use of the Levenberg-Marquardt algorithm, instead of the Gauss-Newton algorithm, to improve the convergence properties of the proposed complex
framework, which in turn reduces the number of layers that redundant calibration pipelines would require. We also compare two previously proposed methods to speedup redundant calibration, namely the alternating direction implicit method and the preconditioned conjugate gradient method. We found that the alternating 
direction implicit method outperforms the preconditioned conjugate gradient method, however the preconditioned conjugate gradient method is more robust as it is more straightforward to adapt to other calibration use cases. 
We mainly focus on the complex framework in this paper rather than the amplitude and phase approach as it makes it trivial to relate the alternating direction implicit method 
and the Levenberg-Marquardt algorithm with each other.
\end{abstract}

@@

Redundant calibration was already quite popular in the 1980's. \citet{Noordam1982} used redundant calibration and the WSRT (Westerbork Synthesis Radio Telescope) to produce high-dynamic range images of 3C84.
The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation, redundant 
calibration is transformed into a linear problem \citep{Wieringa1992,Camps2003,Liu2010}. This approach is commenly referred to as \textsc{logcal}.
We briefly discuss \textsc{logcal} in Appendix~\ref{sec:logcal}. The next improvement in redundant calibration was to perform a firt-order taylor expansion, which is equivalent to the Gauss-Newton (GN) algorithm \citep{Kurien2016}, of the measurement equation an then to solve for the
perturbations of the model-parameters \citep{Liu2010}. The implementation of  this approach is called \textsc{lincal}. \textsc{lincal} is derived from first principals in Appendix~\ref{sec:lincal}. 
\textsc{logcal} and \textsc{lincal} are both implemented in \textsc{omnical}\footnote{https://github.com/jeffzhen/omnical}, which is a layered redundant calibration pipeline \citep{Zheng2014,Ali2015}.
The solutions of \textsc{logcal} are used as initial input for \textsc{lincal}, since the GN algorithm requires that its starting parameter guesses be in the vacinity 
of the solution.

Two ways of speeding up redundant calibration have also been proposed recently. The first technique makes use of the ADI method, in which we alternate between solving the gains, their conjugates and 
the true observed visibilities \citep{Wijnholds2012,Marthi2014}. The second is to use the preconditioned conjugate gradient (PCG) method. PCG allows us to exploit the sparsity inherent in the redundant calibration problem, 
to speed it up \citep{Liu2010}. A proper comparison between these two approaches has not been done before.

New instruments like LOFAR (Low Frequency Array) \citep{Noorishad2012}, PAPER (Precision Array for Probing the Epoch of Reionization) \citep{Ali2015} and HERA (Hydrogen Epoch of Reionization Array) \citep{deboer2015} have recently sparked a renewed interest in redundant calibration. Using redundancy is especially advantageous when we are trying to detect the EoR (Epoch of Reionization) 21 cm signal,
which is inherently quite faint \citep{Parsons2012}. Since redundant baselines make statistically independent measurements of the same $uv$-modes redundancy can greatly 
improve the SNR (signal-to-noise ratio) of EoR detection experiments. This explains why PAPER and HERA, are highly redundant arrays, as they are both primarily used to detect the EoR 21 cm signal.
[GIANNI TO REWRITE AND EXTEND EOR SECTION]

In this paper we propose to use the complex framework introduced by \citet{Smirnov2015} instead of the traditional amplitude and phase approach used by \textsc{lincal}.
We also propose to use the Levenberg-Marquardt gradient-based algorithm instead of the Gauss-Newton algorithm as it exhibits better convergence properties.
In doing so the amount of layers that redundant calibration pipelines need are effectively reduced. We also compare two recent redundant calibration speedup algorithms
with each other, namely ADI \citep{Marthi2014} and PCG \citep{Liu2010}. In this paper, we decided to mainly focus on the complex framework proposed by \citet{Smirnov2015} rather than the traditional phase and amplitude approach as it 
allows us to trivially prove that the ADI method proposed by \citep{Marthi2014} is merely a special case of the more general LM algorithm. Moreover, this framework also 
allows us to show that the ADI method proposed by \citep{Marthi2014} can for all intents and purposes be renamed to redundant \textsc{StEfCal} \citep{Mitchell:MWA-cal,Salvini2014}.

ADD SUMMARY PARAGRAPH WHICH DEPENDS ON PAPER LAYOUT
